---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

# Intro
standard recode manual here: https://www.dhsprogram.com/pubs/pdf/DHSG4/Recode7_DHS_10Sep2018_DHSG4.pdf

# Setup

```{r Setup, message=FALSE}
# Global notebook options
knitr::opts_chunk$set(warning = FALSE, message = FALSE)


#  Run this chunk to load packages and data

rm(list = ls(all = TRUE))

# libraries needed
library(tidyverse)  # most variable creation here uses tidyverse 
library(tidyselect) # used to select variables in FP_EVENTS.R
library(haven)      # used for Haven labeled DHS variables
library(labelled)   # used for Haven labeled variable creation
library(expss)    # for creating tables with Haven labeled data
library(xlsx)     # for exporting to excel
library(naniar)   # to use replace_with_na function
library(here)       # to get R project path
library(sjlabelled) # to set variables label
library(survey)  # to calculate weighted ratio for GAR
library(srvyr)
library(plotly)

#path for R project
here()

# path for this chapter. This is also where the data is stored
chap <- "dhs-raw-data"

# select your survey

# HR Files
HRdatafile <- "ZAHR71DT/ZAHR71FL.DTA"

# PR Files - household member recode
PRdatafile <- "ZAPR71DT/ZAPR71FL.DTA"

# BR Files
BRdatafile <- "ZABR71DT/ZABR71FL.DTA"

# AH files - adult health recode
AHMdatafile <- "ZAAH71DT/ZAAHM71FL.DTA" # men's data
AHWdatafile <- "ZAAH71DT/ZAAHW71FL.DTA" # women's data

# IR files
IRdatafile <- "ZAIR71DT/ZAIR71FL.DTA"

# OD files
ODdatafile <- "ZAOD71DT/ZADV71FL.DTA"

# CR files
CRdatafile <- "ZACR71DT/ZACR71FL.DTA"

# MR files
MRdatafile <- "ZAMR71DT/ZAMR71FL.DTA"

# ****************************

# function to print exponentiated coefficients and confidence interals

exponentiated_output <- function(model) {
  output <- exp(cbind(OR = coefficients(model), confint(model)))
  print(output)
}

```

``` {r Load datasets}
# open datasets
# HRdata <-  read_dta(here(chap, HRdatafile))

PRdata <-  read_dta(here(chap, PRdatafile))

# BRdata <-  read_dta(here(chap, BRdatafile))

AHMdata <- read_dta(here(chap, AHMdatafile))

AHWdata <- read_dta(here(chap, AHWdatafile))

# IRdata <- read_dta(here(chap, IRdatafile))

# ODdata <- read_dta(here(chap, ODdatafile))

# CRdata <- read_dta(here(chap, CRdatafile))

# MRdata <- read_dta(here(chap, MRdatafile))

```


# Alcohol

## Men

```{r}

# view(get_label(AHMdata))

# The below code confirms that missing values for sm919 should be treated as zeroes. Those who responded 'no' to sm916 (ever consumed alcohol) were not asked sm917 (consumed alcohol in last 12m). Those who responded 'no' to the latter were not asked sm919
# AHMdata %>%
#   select(
#     mv012, #age
#     mv024, #province
#     sm916:sm924
#   ) %>%
#   mutate(drinks_wk = sm919a + sm919b + sm919c + sm919d + sm919e + sm919f + sm919g) %>%
#   mutate(drinks_last_wk_recorded = ifelse(!is.na(drinks_wk), 1, NA)) %>%
#   count(sm916, sm917, drinks_last_wk_recorded) %>%
#   view()

etoh_m <- AHMdata %>% 
  select(
    mv012, #age
    mv024, #province
    sm919a, sm919b, sm919c, sm919d, sm919e, sm919f, sm919g
  ) %>%
  mutate(province = as_character(mv024)) %>% 
  # mutate(missing_check = rowSums(is.na(select(., starts_with("sm919"))))) %>% # sum of missing values across drinks per day columns
  # mutate(drinking_freq = rowSums(select(., starts_with("sm919") != 0))) %>% # sum of days drinking across drinks per day columns
  mutate(drinks_wk = sm919a + sm919b + sm919c + sm919d + sm919e + sm919f + sm919g) %>% 
  mutate(drinks_wk = ifelse(is.na(drinks_wk), 0, drinks_wk)) %>% 
  mutate(etoh_5 = ifelse(if_any(sm919a:sm919g, ~.x >= 5), 1, 0)) %>% 
  mutate(etoh_5 = ifelse(is.na(etoh_5), 0, etoh_5)) %>% 
  mutate(anyalc = ifelse(drinks_wk > 0, 1, 0)) %>% 
  # mutate(average_dd = drinks_wk / drinkingfreq) %>% # not used currently
  mutate(highalc = ifelse(drinks_wk > 23, 1, 0)) %>% 
  mutate(highalc = ifelse(is.na(highalc), 0, highalc)) %>% #Missing values should be interpreted as no alcohol intake
  rename(
    age = mv012
  ) %>% 
  mutate(age2 = age^2, age3 = age^3) %>% 
  select(age, age2, age3, province, highalc, anyalc, drinks_wk, etoh_5)

```

### Testing ordinal LR

```{r}


drinking_levels <- c("none", "1-10", "11-20", ">20")

etoh_cat_m <- etoh_m %>%
  mutate(etohcat = ifelse(drinks_wk == 0, "none",
                          ifelse(drinks_wk >0 & drinks_wk <=10, "1-10",
                                 ifelse(drinks_wk > 10 & drinks_wk <=20, "11-20",
                                        ifelse(drinks_wk >20, ">20", NA))))) %>%
  mutate(etohcat = factor(etohcat, levels = drinking_levels))

etoh_cat_model <- MASS::polr(
  etohcat ~ age + scale(age2) + province, # needed to scale age2 to prevent NaNs from output
  data = etoh_cat_m,
  Hess = TRUE
)

summary(etoh_cat_model)

exponentiated_output(etoh_cat_model)

```

### Model highalc ~ age

``` {r}
etoh_model_m <- glm(
  formula = highalc ~ age + age2 + age3,
  family = binomial(link = "logit"),
  data = etoh_m
)

exponentiated_output(etoh_model_m)

```

### Model highalc ~ age + province (cubic function)

``` {r}

etoh_model_m <- glm(
  formula = highalc ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = etoh_m
)

etoh_coeffs_m <- exp(coefficients(etoh_model_m))

exponentiated_output(etoh_model_m)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_m[5:12]) #1 is for ref group EC

etoh_prev_m_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "alcohol", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_m[1] * etoh_coeffs_m[2]^age * etoh_coeffs_m[3]^(age^2) * etoh_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_m_c <- bind_rows(etoh_prev_m_c, temp)
}

etoh_prev_m_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model alcohol ~ age + province (quadratic function)

``` {r}

etoh_model_m <- glm(
  formula = highalc ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = etoh_m
)

etoh_coeffs_m <- exp(coefficients(etoh_model_m))

summary(etoh_model_m)

exponentiated_output(etoh_model_m)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_m[4:11]) #1 is for ref group EC

etoh_prev_m_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "alcohol", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_m[1] * etoh_coeffs_m[2]^age * etoh_coeffs_m[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_m_q <- bind_rows(etoh_prev_m_q, temp)
}

etoh_prev_m_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model anyalc ~ age + province (cubic function)

``` {r}

etoh_model_m <- glm(
  formula = anyalc ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = etoh_m
)

etoh_coeffs_m <- exp(coefficients(etoh_model_m))

summary(etoh_model_m)

exponentiated_output(etoh_model_m)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_m[5:12]) #1 is for ref group EC

etoh_prev_m_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "alcohol", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_m[1] * etoh_coeffs_m[2]^age * etoh_coeffs_m[3]^(age^2) * etoh_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_m_c <- bind_rows(etoh_prev_m_c, temp)
}

etoh_prev_m_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model anyalc ~ age + province (quadratic function)

``` {r}

etoh_model_m <- glm(
  formula = anyalc ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = etoh_m
)

etoh_coeffs_m <- exp(coefficients(etoh_model_m))

summary(etoh_model_m)

exponentiated_output(etoh_model_m)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_m[4:11]) #1 is for ref group EC

etoh_prev_m_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "alcohol", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_m[1] * etoh_coeffs_m[2]^age * etoh_coeffs_m[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_m_q <- bind_rows(etoh_prev_m_q, temp)
}

etoh_prev_m_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model etoh_5 ~ age + province (quadratic function)

``` {r}

etoh_model_m <- glm(
  formula = etoh_5 ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = etoh_m
)

etoh_coeffs_m <- exp(coefficients(etoh_model_m))

summary(etoh_model_m)

exponentiated_output(etoh_model_m)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_m[4:11]) #1 is for ref group EC

etoh_prev_m_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "alcohol", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_m[1] * etoh_coeffs_m[2]^age * etoh_coeffs_m[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_m_q <- bind_rows(etoh_prev_m_q, temp)
}

etoh_prev_m_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

# TESTING PREDICT() FUNCTION
# This works - can refactor code later for all other predictions
# d <- data.frame(
#   age = rep(15:90, each = 9),
#   age2 = rep((15:90)^2, each = 9),
#   province = unique(etoh_m$province)
# )
# 
# predict_test <- cbind(
#   d, 
#   prevalence = predict(etoh_model_m, d, type = "response"),
#   risk_factor = "alcohol"
# )
# 
# predict_test %>% 
#   plot_ly(
#     x = ~age,
#     y = ~prevalence,
#     color = ~province,
#     mode = "lines"
#   )

```

## Women

```{r}

# view(get_label(AHWdata))

# The below code confirms that missing values for s1227 should be treated as zeroes. Those who responded 'no' to s1224 (ever consumed alcohol) were not asked s1225 (consumed alcohol in last 12m). Those who responded 'no' to the latter were not asked s1227
# AHWdata %>%
#   select(
#     v012, #age
#     v024, #province
#     s1224:s1228
#   ) %>%
#   mutate(drinks_wk = s1227a + s1227b + s1227c + s1227d + s1227e + s1227f + s1227g) %>%
#   mutate(drinks_last_wk_recorded = ifelse(!is.na(drinks_wk), 1, NA)) %>%
#   count(s1224, s1225, drinks_last_wk_recorded) %>%
#   view()

etoh_w <- AHWdata %>% 
  select(
    v012, #age
    v024, #province
    s1227a, s1227b, s1227c, s1227d, s1227e, s1227f, s1227g
  ) %>% 
  mutate(province = as_character(v024)) %>% 
  # mutate(drinkingfreq = rowSums(select(., starts_with("s1227")) != 0)) %>% 
  mutate(drinks_wk = s1227a + s1227b + s1227c + s1227d + s1227e + s1227f + s1227g) %>% 
  mutate(drinks_wk = ifelse(is.na(drinks_wk), 0, drinks_wk)) %>% 
  mutate(etoh_5 = ifelse(if_any(s1227a:s1227g, ~.x >= 5), 1, 0)) %>% 
  mutate(etoh_5 = ifelse(is.na(etoh_5), 0, etoh_5)) %>% 
  mutate(anyalc = ifelse(drinks_wk > 0, 1, 0)) %>% 
  # mutate(average_dd = drinks_wk / drinkingfreq) %>% 
  mutate(highalc = ifelse(drinks_wk > 23, 1, 0)) %>% 
  mutate(highalc = ifelse(is.na(highalc), 0, highalc)) %>% #Missing values should be interpreted as no alcohol intake
  rename(
    age = v012
  ) %>% 
  mutate(age2 = age^2, age3 = age^3) %>% 
  select(age, age2, age3, province, highalc, anyalc, etoh_5)
```

### Model highalc ~ age
``` {r}
etoh_model_w <- glm(
  formula = highalc ~ age + age2 + age3,
  family = binomial(),
  data = etoh_w
)

exponentiated_output(etoh_model_w)

```

### Model highalc ~ age + province (cubic function)

``` {r}

etoh_model_w <- glm(
  formula = highalc ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = etoh_w
)

etoh_coeffs_w <- exp(coefficients(etoh_model_w))

exponentiated_output(etoh_model_w)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_w[5:12]) #1 is for ref group EC

etoh_prev_w_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "alcohol", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_w[1] * etoh_coeffs_w[2]^age * etoh_coeffs_w[3]^(age^2) * etoh_coeffs_w[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_w_c <- bind_rows(etoh_prev_w_c, temp)
}

etoh_prev_w_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model highalc ~ age + province (quadratic function)

``` {r}

etoh_model_w <- glm(
  formula = highalc ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = etoh_w
)

etoh_coeffs_w <- exp(coefficients(etoh_model_w))

exponentiated_output(etoh_model_w)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_w[4:11]) #1 is for ref group EC

etoh_prev_w_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "alcohol", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_w[1] * etoh_coeffs_w[2]^age * etoh_coeffs_w[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_w_q <- bind_rows(etoh_prev_w_q, temp)
}

etoh_prev_w_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model anyalc ~ age + province (cubic function)

``` {r}

etoh_model_w <- glm(
  formula = anyalc ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = etoh_w
)

etoh_coeffs_w <- exp(coefficients(etoh_model_w))

summary(etoh_model_w)

exponentiated_output(etoh_model_w)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_w[5:12]) #1 is for ref group EC

etoh_prev_w_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "alcohol", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_w[1] * etoh_coeffs_w[2]^age * etoh_coeffs_w[3]^(age^2) * etoh_coeffs_w[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_w_c <- bind_rows(etoh_prev_w_c, temp)
}

etoh_prev_w_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model anyalc ~ age + province (quadratic function)

``` {r}

etoh_model_w <- glm(
  formula = anyalc ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = etoh_w
)

etoh_coeffs_w <- exp(coefficients(etoh_model_w))

summary(etoh_model_w)

exponentiated_output(etoh_model_w)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_w[4:11]) #1 is for ref group EC

etoh_prev_w_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "alcohol", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_w[1] * etoh_coeffs_w[2]^age * etoh_coeffs_w[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_w_q <- bind_rows(etoh_prev_w_q, temp)
}

etoh_prev_w_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

### Model etoh_5 ~ age + province (quadratic function)

``` {r}

etoh_model_w <- glm(
  formula = etoh_5 ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = etoh_w
)

etoh_coeffs_w <- exp(coefficients(etoh_model_w))

summary(etoh_model_w)

exponentiated_output(etoh_model_w)

```

#### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, etoh_coeffs_w[4:11]) #1 is for ref group EC

etoh_prev_w_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "alcohol", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = etoh_coeffs_w[1] * etoh_coeffs_w[2]^age * etoh_coeffs_w[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(
      prevalence = odds / (1 + odds)
    ) %>% 
    select(-odds)
  
  etoh_prev_w_q <- bind_rows(etoh_prev_w_q, temp)
}

etoh_prev_w_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

# Smoking

Leigh uses `mv463aa` as the smoking variable, where a value of 1 = smoking every day, and 2 = smoking some days.

```{r}

# view(get_label(AHMdata))

# view(get_label(AHWdata))

# AHMdata %>%
#   select(mv463aa) %>% 
#   get_labels()

```

## Men

### Wrangle

```{r}

smoking_m <- AHMdata %>% 
  select(
    mv012, # age
    mv024, # province
    mv463aa
  ) %>% 
  mutate(province = as_character(mv024)) %>% 
  mutate(
    smoking = ifelse(mv463aa == 1 | mv463aa ==2, 1, 0)
  ) %>% 
  rename(age = mv012) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3 
  ) %>% 
  select(-mv463aa) %>% 
  relocate(age, age2, age3, province, smoking)

```

### Model smoking ~ age

``` {r}

smoking_model_m <- glm(
  formula = smoking ~ age + age2 + age3,
  family = binomial(link = "logit"),
  data = smoking_m
)

exponentiated_output(smoking_model_m)

```

### Model smoking ~ age + province

``` {r}

smoking_model_m <- glm(
  formula = smoking ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = smoking_m
)

smoking_coeffs_m <- exp(coefficients(smoking_model_m))

exponentiated_output(smoking_model_m)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, smoking_coeffs_m[5:12]) #1 is for ref group EC

smoking_prev_m_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "smoking", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = smoking_coeffs_m[1] * smoking_coeffs_m[2]^age * smoking_coeffs_m[3]^(age^2) * smoking_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  smoking_prev_m_c <- bind_rows(smoking_prev_m_c, temp)
}

smoking_prev_m_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )

```

## Women

### Wrangle

```{r}

smoking_f <- AHWdata %>% 
  select(
    v012, # age
    v024, # province
    v463aa
  ) %>% 
  mutate(province = as_character(v024)) %>% 
  mutate(
    smoking = ifelse(v463aa == 1 | v463aa ==2, 1, 0)
  ) %>% 
  rename(age = v012) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3 
  ) %>% 
  select(-v463aa) %>% 
  relocate(age, age2, age3, province, smoking)

```

### Model smoking ~ age

``` {r}

smoking_model_f <- glm(
  formula = smoking ~ age + age2 + age3,
  family = binomial(link = "logit"),
  data = smoking_f
)

exponentiated_output(smoking_model_f)

```

### Model smoking ~ age + province

``` {r}

smoking_model_f <- glm(
  formula = smoking ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = smoking_f
)

smoking_coeffs_f <- exp(coefficients(smoking_model_f))

exponentiated_output(smoking_model_f)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, smoking_coeffs_f[5:12]) #1 is for ref group EC

smoking_prev_f_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "smoking", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = smoking_coeffs_f[1] * smoking_coeffs_f[2]^age * smoking_coeffs_f[3]^(age^2) * smoking_coeffs_f[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  smoking_prev_f_c <- bind_rows(smoking_prev_f_c, temp)
}

# smoking_prev_f %>% 
#   ggplot(aes(x = age, y = prevalence, color = province)) +
#   geom_line()

smoking_prev_f_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  ) 

```

# Underweight

## View PR data variables
```{r}
# PRdata %>% 
#   get_label() %>% 
#   view()
```

## Wrangle
```{r}

bmi_both <- PRdata %>% 
  select(
    hv105, #age
    hv104, #sex
    hv024, #province
    ha3, hb3, #height women's and men's in cm
    ha2, hb2 #weight women's and men's 
  ) %>%
  mutate(province = as_character(hv024)) %>% 
  rename(
    age = hv105,
    sex = hv104,
    ht_w = ha3,
    ht_m = hb3,
    wt_w = ha2,
    wt_m = hb2
  ) %>% 
  mutate(
    bmi = ifelse(
      sex == 1 & ht_m < 2500 & wt_m < 2500 & !is.na(ht_m) & !is.na(wt_m),
      (wt_m / 10) / (ht_m / 1000)^2,
      ifelse(
        sex == 2 & ht_w < 2500 & wt_w < 2500 & !is.na(ht_w) & !is.na(wt_w),
        (wt_w / 10) / (ht_w / 1000)^2,
        NA
      )
    )
  ) %>% 
  mutate(
    underwt = ifelse(
      bmi < 18.5, 1, 0
    )
  ) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3
  ) %>% 
  select(sex, province, age, age2, age3, underwt)

```

## Males

``` {r}
underwt_m <- bmi_both %>% 
  filter(sex == 1)
```

### underweight ~ age

```{r}
underwt_model_m <- glm(
  formula = underwt ~ age + age2 + age3,
  family = binomial(),
  data = underwt_m
)

exponentiated_output(underwt_model_m)

```

### Model underwt ~ age + province

``` {r}

underwt_model_m <- glm(
  formula = underwt ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = underwt_m
)

underwt_coeffs_m <- exp(coefficients(underwt_model_m))

exponentiated_output(underwt_model_m)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, underwt_coeffs_m[5:12]) #1 is for ref group EC

underwt_prev_m_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "underweight", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = underwt_coeffs_m[1] * underwt_coeffs_m[2]^age * underwt_coeffs_m[3]^(age^2) * underwt_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  underwt_prev_m_c <- bind_rows(underwt_prev_m_c, temp)
}

# underwt_prev_m %>% 
#   ggplot(aes(x = age, y = prevalence, color = province)) +
#   geom_line()

underwt_prev_m_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )
```


## Females

``` {r}
underwt_f <- bmi_both %>% 
  filter(sex == 2)
```

### underweight ~ age

``` {r}
underwt_model_f <- glm(
  formula = underwt ~ age + age2 + age3,
  family = binomial(),
  data = underwt_f
)

exponentiated_output(underwt_model_f)

```

### Model underwt ~ age + province

``` {r}

underwt_model_f <- glm(
  formula = underwt ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = underwt_f
)

underwt_coeffs_f <- exp(coefficients(underwt_model_f))

exponentiated_output(underwt_model_f)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, underwt_coeffs_f[5:12]) #1 is for ref group EC

underwt_prev_f_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "underweight", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = underwt_coeffs_f[1] * underwt_coeffs_f[2]^age * underwt_coeffs_f[3]^(age^2) * underwt_coeffs_f[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  underwt_prev_f_c <- bind_rows(underwt_prev_f_c, temp)
}

# underwt_prev_f %>% 
#   ggplot(aes(x = age, y = prevalence, color = province)) +
#   geom_line()

underwt_prev_f_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  )
```


# Diabetes

## Wrangle

```{r}

diabetes <- PRdata %>% 
  select(
    hv104, #sex
    hv105, #age
    hv024, #province
    shmhba1c,
    shwhba1c
  ) %>% 
  mutate(
    province = as_character(hv024)
  ) %>% 
  mutate(
    hba1c = ifelse(
      hv104 == 1, 
      shmhba1c,
      ifelse(
        hv104 == 2,
        shwhba1c,
        NA
      )
    )
  ) %>% 
  rename(
    sex = hv104,
    age = hv105
  ) %>% 
  mutate(
    adjhba1c = (hba1c - 228) / 0.9866
  ) %>% 
  mutate(
    diabetes = ifelse(adjhba1c >= 6500, 1, 0)
  ) %>% 
  select(sex, age, province, diabetes) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3
  )
```

## Men
``` {r}
diabetes_m <- diabetes %>% 
  filter(sex == 1)
```

### Model diabetes ~ age
``` {r}
dm_model_m <- glm(
  formula = diabetes ~ age + age2 + age3,
  family = binomial(),
  data = diabetes_m
)

exponentiated_output(dm_model_m)
```

### Model diabetes ~ age + province (cubic function)

``` {r}

dm_model_m <- glm(
  formula = diabetes ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = diabetes_m
)

dm_coeffs_m <- exp(coefficients(dm_model_m))

exponentiated_output(dm_model_m)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, dm_coeffs_m[5:12]) #1 is for ref group EC

dm_prev_m_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "diabetes", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = dm_coeffs_m[1] * dm_coeffs_m[2]^age * dm_coeffs_m[3]^(age^2) * dm_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  dm_prev_m_c <- bind_rows(dm_prev_m_c, temp)
}

dm_prev_m_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  ) 

```

### Model diabetes ~ age + province (quadratic function)

``` {r}

dm_model_m <- glm(
  formula = diabetes ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = diabetes_m
)

dm_coeffs_m <- exp(coefficients(dm_model_m))

exponentiated_output(dm_model_m)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, dm_coeffs_m[4:11]) #1 is for ref group EC

dm_prev_m_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "diabetes", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = dm_coeffs_m[1] * dm_coeffs_m[2]^age * dm_coeffs_m[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  dm_prev_m_q <- bind_rows(dm_prev_m_q, temp)
}

dm_prev_m_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  ) 

```
## Women
``` {r}
diabetes_f <- diabetes %>% 
  filter(sex == 2)
```

### Model diabetes ~ age
``` {r}
dm_model_f <- glm(
  formula = diabetes ~ age + age2 + age3,
  family = binomial(),
  data = diabetes_f
)

exponentiated_output(dm_model_f)

```


### Model diabetes ~ age + province (cubic function)

``` {r}

dm_model_f <- glm(
  formula = diabetes ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = diabetes_f
)

dm_coeffs_f <- exp(coefficients(dm_model_f))

exponentiated_output(dm_model_f)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, dm_coeffs_f[5:12]) #1 is for ref group EC

dm_prev_f_c <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "diabetes", model = "cubic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = dm_coeffs_f[1] * dm_coeffs_f[2]^age * dm_coeffs_f[3]^(age^2) * dm_coeffs_f[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  dm_prev_f_c <- bind_rows(dm_prev_f_c, temp)
}

dm_prev_f_c %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  ) 

```

### Model diabetes ~ age + province (quadratic function)

``` {r}

dm_model_f <- glm(
  formula = diabetes ~ age + age2 + province,
  family = binomial(link = "logit"),
  data = diabetes_f
)

dm_coeffs_f <- exp(coefficients(dm_model_f))

exponentiated_output(dm_model_f)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, dm_coeffs_f[4:11]) #1 is for ref group EC

dm_prev_f_q <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "diabetes", model = "quadratic")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = dm_coeffs_f[1] * dm_coeffs_f[2]^age * dm_coeffs_f[3]^(age^2) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  dm_prev_f_q <- bind_rows(dm_prev_f_q, temp)
}

dm_prev_f_q %>% 
  plot_ly(
    x = ~age,
    y = ~prevalence,
    color = ~province,
    mode = "lines"
  ) 

```

# Combine outputs

```{r}
# output <- bind_rows(
#   etoh_prev_m_c, etoh_prev_m_q,
#   etoh_prev_w_c, etoh_prev_w_q,
#   smoking_prev_m_c, smoking_prev_f_c,
#   underwt_prev_f_c, underwt_prev_m_c,
#   dm_prev_m_c, dm_prev_m_q, 
#   dm_prev_f_c, dm_prev_f_q
# ) %>% 
#   pivot_wider(names_from = c(risk_factor, sex), values_from = prevalence)

# openxlsx::write.xlsx(output, "provincial_estimates.xlsx")

```


