---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Intro
standard recode manual here: https://www.dhsprogram.com/pubs/pdf/DHSG4/Recode7_DHS_10Sep2018_DHSG4.pdf

# Setup

```{r Setup, message=FALSE}
#  Run this chunk to load packages and data

rm(list = ls(all = TRUE))

# libraries needed
library(tidyverse)  # most variable creation here uses tidyverse 
library(tidyselect) # used to select variables in FP_EVENTS.R
library(haven)      # used for Haven labeled DHS variables
library(labelled)   # used for Haven labeled variable creation
library(expss)    # for creating tables with Haven labeled data
library(xlsx)     # for exporting to excel
library(naniar)   # to use replace_with_na function
library(here)       # to get R project path
library(sjlabelled) # to set variables label
library(survey)  # to calculate weighted ratio for GAR
library(srvyr)

#path for R project
here()

# path for this chapter. This is also where the data is stored
chap <- "dhs-raw-data"

# select your survey

# HR Files
HRdatafile <- "ZAHR71DT/ZAHR71FL.DTA"

# PR Files - household member recode
PRdatafile <- "ZAPR71DT/ZAPR71FL.DTA"

# BR Files
BRdatafile <- "ZABR71DT/ZABR71FL.DTA"

# AH files - adult health recode
AHMdatafile <- "ZAAH71DT/ZAAHM71FL.DTA" # men's data
AHWdatafile <- "ZAAH71DT/ZAAHW71FL.DTA" # women's data

# IR files
IRdatafile <- "ZAIR71DT/ZAIR71FL.DTA"

# OD files
ODdatafile <- "ZAOD71DT/ZADV71FL.DTA"

# CR files
CRdatafile <- "ZACR71DT/ZACR71FL.DTA"

# MR files
MRdatafile <- "ZAMR71DT/ZAMR71FL.DTA"

# ****************************

```

``` {r Load datasets}
# open datasets
# HRdata <-  read_dta(here(chap, HRdatafile))

PRdata <-  read_dta(here(chap, PRdatafile))

# BRdata <-  read_dta(here(chap, BRdatafile))

AHMdata <- read_dta(here(chap, AHMdatafile))

AHWdata <- read_dta(here(chap, AHWdatafile))

# IRdata <- read_dta(here(chap, IRdatafile))

# ODdata <- read_dta(here(chap, ODdatafile))

# CRdata <- read_dta(here(chap, CRdatafile))

# MRdata <- read_dta(here(chap, MRdatafile))

```


# Alcohol

## Men

```{r}

# view(get_label(AHMdata))

etoh_m <- AHMdata %>% 
  select(
    mv012, #age
    mv005, #sample weight
    sm919a, sm919b, sm919c, sm919d, sm919e, sm919f, sm919g
  ) %>%
  mutate(wt = mv005 / 1000000) %>% 
  mutate(drinkingfreq = rowSums(select(., starts_with("sm919")) != 0)) %>% 
  mutate(drinks_wk = sm919a + sm919b + sm919c + sm919d + sm919e + sm919f + sm919g) %>% 
  mutate(average_dd = drinks_wk / drinkingfreq) %>% 
  mutate(highalc = ifelse(drinks_wk > 23, 1, 0)) %>% 
  rename(
    age = mv012
  ) %>% 
  mutate(age2 = age^2, age3 = age^3) %>% 
  select(age, age2, age3, highalc, wt)

etoh_m_model <- glm(
  formula = highalc ~ age + age2 + age3,
  family = binomial(),
  data = etoh_m
)

exp(coefficients(etoh_m_model))

```

## Women

```{r}

# view(get_label(AHWdata))

etoh_w <- AHWdata %>% 
  select(
    v012, #age
    s1227a, s1227b, s1227c, s1227d, s1227e, s1227f, s1227g
  ) %>% 
  mutate(drinkingfreq = rowSums(select(., starts_with("s1227")) != 0)) %>% 
  mutate(drinks_wk = s1227a + s1227b + s1227c + s1227d + s1227e + s1227f + s1227g) %>% 
  mutate(average_dd = drinks_wk / drinkingfreq) %>% 
  mutate(highalc = ifelse(drinks_wk > 23, 1, 0)) %>% 
  rename(
    age = v012
  ) %>% 
  mutate(age2 = age^2, age3 = age^3) %>% 
  select(age, age2, age3, highalc)

etoh_w_model <- glm(
  formula = highalc ~ age + age2 + age3,
  family = binomial(),
  data = etoh_w
)

exp(coefficients(etoh_w_model))

```

# Smoking

Leigh uses `mv463aa` as the smoking variable, where a value of 1 = smoking every day, and 2 = smoking some days.

```{r}

# view(get_label(AHMdata))

# view(get_label(AHWdata))

# AHMdata %>%
#   select(mv463aa) %>% 
#   get_labels()

```

## Men

### Wrangle

```{r}

smoking_m <- AHMdata %>% 
  select(
    mv012, # age
    mv024, # province
    mv463aa
  ) %>% 
  mutate(province = as_character(mv024)) %>% 
  mutate(
    smoking = ifelse(mv463aa == 1 | mv463aa ==2, 1, 0)
  ) %>% 
  rename(age = mv012) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3 
  ) %>% 
  select(-mv463aa) %>% 
  relocate(age, age2, age3, province, smoking)

```

### Model smoking ~ age

``` {r}

smoking_model_m <- glm(
  formula = smoking ~ age + age2 + age3,
  family = binomial(link = "logit"),
  data = smoking_m
)

exp(coefficients(smoking_model_m))

```

### Model smoking ~ age + province

``` {r}

smoking_model_m <- glm(
  formula = smoking ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = smoking_m
)

smoking_coeffs_m <- exp(coefficients(smoking_model_m))

summary(smoking_model_m)
```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, smoking_coeffs_m[5:12]) #1 is for ref group EC

smoking_prev_m <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "smoking")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = smoking_coeffs_m[1] * smoking_coeffs_m[2]^age * smoking_coeffs_m[3]^(age^2) * smoking_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  smoking_prev_m <- bind_rows(smoking_prev_m, temp)
}

smoking_prev_m %>% 
  ggplot(aes(x = age, y = prevalence, color = province)) + 
  geom_line()

```

## Women

### Wrangle

```{r}

smoking_f <- AHWdata %>% 
  select(
    v012, # age
    v024, # province
    v463aa
  ) %>% 
  mutate(province = as_character(v024)) %>% 
  mutate(
    smoking = ifelse(v463aa == 1 | v463aa ==2, 1, 0)
  ) %>% 
  rename(age = v012) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3 
  ) %>% 
  select(-v463aa) %>% 
  relocate(age, age2, age3, province, smoking)

```

### Model smoking ~ age

``` {r}

smoking_model_f <- glm(
  formula = smoking ~ age + age2 + age3,
  family = binomial(link = "logit"),
  data = smoking_f
)

summary(smoking_model_f)

smoking_coeffs_f <- exp(coefficients(smoking_model_f))

exp(confint(smoking_model_f))

```

### Model smoking ~ age + province

``` {r}

smoking_model_f <- glm(
  formula = smoking ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = smoking_f
)

smoking_coeffs_f <- exp(coefficients(smoking_model_f))

summary(smoking_model_f)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, smoking_coeffs_f[5:12]) #1 is for ref group EC

smoking_prev_f <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "smoking")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = smoking_coeffs_f[1] * smoking_coeffs_f[2]^age * smoking_coeffs_f[3]^(age^2) * smoking_coeffs_f[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  smoking_prev_f <- bind_rows(smoking_prev_f, temp)
}

smoking_prev_f %>% 
  ggplot(aes(x = age, y = prevalence, color = province)) +
  geom_line()

```

# Underweight

## View PR data variables
```{r}
PRdata %>% 
  get_label() %>% 
  view()
```

## Wrangle
```{r}

bmi_both <- PRdata %>% 
  select(
    hv105, #age
    hv104, #sex
    hv024, #province
    ha3, hb3, #height women's and men's in cm
    ha2, hb2 #weight women's and men's 
  ) %>%
  mutate(province = as_character(hv024)) %>% 
  rename(
    age = hv105,
    sex = hv104,
    ht_w = ha3,
    ht_m = hb3,
    wt_w = ha2,
    wt_m = hb2
  ) %>% 
  mutate(
    bmi = ifelse(
      sex == 1 & ht_m < 2500 & wt_m < 2500 & !is.na(ht_m) & !is.na(wt_m),
      (wt_m / 10) / (ht_m / 1000)^2,
      ifelse(
        sex == 2 & ht_w < 2500 & wt_w < 2500 & !is.na(ht_w) & !is.na(wt_w),
        (wt_w / 10) / (ht_w / 1000)^2,
        NA
      )
    )
  ) %>% 
  mutate(
    underwt = ifelse(
      bmi < 18.5, 1, 0
    )
  ) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3
  ) %>% 
  select(sex, province, age, age2, age3, underwt)

```

## Males

``` {r}
underwt_m <- bmi_both %>% 
  filter(sex == 1)
```

### underweight ~ age

```{r}
underwt_model_m <- glm(
  formula = underwt ~ age + age2 + age3,
  family = binomial(),
  data = underwt_m
)

exp(coefficients(underwt_model_m))

```

### Model underwt ~ age + province

``` {r}

underwt_model_m <- glm(
  formula = underwt ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = underwt_m
)

underwt_coeffs_m <- exp(coefficients(underwt_model_m))

summary(underwt_model_m)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, underwt_coeffs_m[5:12]) #1 is for ref group EC

underwt_prev_m <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "underweight")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = underwt_coeffs_m[1] * underwt_coeffs_m[2]^age * underwt_coeffs_m[3]^(age^2) * underwt_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  underwt_prev_m <- bind_rows(underwt_prev_m, temp)
}

underwt_prev_m %>% 
  ggplot(aes(x = age, y = prevalence, color = province)) +
  geom_line()

```


## Females

``` {r}
underwt_f <- bmi_both %>% 
  filter(sex == 2)
```

### underweight ~ age

``` {r}
underwt_model_f <- glm(
  formula = underwt ~ age + age2 + age3,
  family = binomial(),
  data = underwt_f
)

exp(coefficients(underwt_model_f))


```

### Model underwt ~ age + province

``` {r}

underwt_model_f <- glm(
  formula = underwt ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = underwt_f
)

underwt_coeffs_f <- exp(coefficients(underwt_model_f))

summary(underwt_model_f)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, underwt_coeffs_f[5:12]) #1 is for ref group EC

underwt_prev_f <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "underweight")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = underwt_coeffs_f[1] * underwt_coeffs_f[2]^age * underwt_coeffs_f[3]^(age^2) * underwt_coeffs_f[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  underwt_prev_f <- bind_rows(underwt_prev_f, temp)
}

underwt_prev_f %>% 
  ggplot(aes(x = age, y = prevalence, color = province)) +
  geom_line()

```


# Diabetes

## Wrangle

```{r}

diabetes <- PRdata %>% 
  select(
    hv104, #sex
    hv105, #age
    hv024, #province
    shmhba1c,
    shwhba1c
  ) %>% 
  mutate(
    province = as_character(hv024)
  ) %>% 
  mutate(
    hba1c = ifelse(
      hv104 == 1, 
      shmhba1c,
      ifelse(
        hv104 == 2,
        shwhba1c,
        NA
      )
    )
  ) %>% 
  rename(
    sex = hv104,
    age = hv105
  ) %>% 
  mutate(
    adjhba1c = (hba1c - 228) / 0.9866
  ) %>% 
  mutate(
    diabetes = ifelse(adjhba1c >= 6500, 1, 0)
  ) %>% 
  select(sex, age, province, diabetes) %>% 
  mutate(
    age2 = age^2,
    age3 = age^3
  )
```

## Men
``` {r}
diabetes_m <- diabetes %>% 
  filter(sex == 1)
```

### Model diabetes ~ age
``` {r}
dm_model_m <- glm(
  formula = diabetes ~ age + age2 + age3,
  family = binomial(),
  data = diabetes_m
)

exp(coefficients(dm_model_m))
```

### Model diabetes ~ age + province

``` {r}

dm_model_m <- glm(
  formula = diabetes ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = diabetes_m
)

dm_coeffs_m <- exp(coefficients(dm_model_m))

summary(dm_model_m)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, dm_coeffs_m[5:12]) #1 is for ref group EC

dm_prev_m <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "males", risk_factor = "diabetes")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = dm_coeffs_m[1] * dm_coeffs_m[2]^age * dm_coeffs_m[3]^(age^2) * dm_coeffs_m[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  dm_prev_m <- bind_rows(dm_prev_m, temp)
}

dm_prev_m %>% 
  ggplot(aes(x = age, y = prevalence, color = province)) +
  geom_line()

```

## Women
``` {r}
diabetes_f <- diabetes %>% 
  filter(sex == 2)
```

### Model diabetes ~ age
``` {r}
dm_model_f <- glm(
  formula = diabetes ~ age + age2 + age3,
  family = binomial(),
  data = diabetes_f
)

exp(coefficients(dm_model_f))

```


### Model diabetes ~ age + province

``` {r}

dm_model_f <- glm(
  formula = diabetes ~ age + age2 + age3 + province,
  family = binomial(link = "logit"),
  data = diabetes_f
)

dm_coeffs_f <- exp(coefficients(dm_model_f))

summary(dm_model_f)

```

### Prevalence outputs

```{r}

age <- c(15:90)

provinces <- c("EC", "FS", "GP", "KZN", "LP", "MP", "NW", "NC", "WC")

provincial_coeffs <- c(1, dm_coeffs_f[5:12]) #1 is for ref group EC

dm_prev_f <- tibble()

for (n in 1:9) {
  temp <- tibble(age = age, sex = "females", risk_factor = "diabetes")
  
  temp <- temp %>% 
    mutate(province = provinces[n]) %>% 
    mutate(odds = dm_coeffs_f[1] * dm_coeffs_f[2]^age * dm_coeffs_f[3]^(age^2) * dm_coeffs_f[4]^(age^3) * provincial_coeffs[n]) %>% 
    mutate(prevalence = odds / (1 + odds)) %>% 
    select(-odds)
  
  dm_prev_f <- bind_rows(dm_prev_f, temp)
}

dm_prev_m %>% 
  ggplot(aes(x = age, y = prevalence, color = province)) +
  geom_line()

```


# Combine outputs

```{r}
output <- bind_rows(
  smoking_prev_f, smoking_prev_m,
  underwt_prev_f, underwt_prev_m,
  dm_prev_f, dm_prev_m
) %>% 
  pivot_wider(names_from = c(risk_factor, sex), values_from = prevalence)

# write.xlsx(output, file = "provincial_estimates.xlsx", row.names = FALSE)

```


